---
title: "Temperature and Freshwater as Methods for Controlling *D. Vexillum* Biofouling"
author: "Payton Arthur, Hazel de Haas, Lauren Gill"
date: "27/11/2021"
  
output:
  pdf_document:
    toc: TRUE
toc-depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, tidy=TRUE, tidy.opts = list(width.cutoff = 70))
```

## Abstract

Biofouling, the unwanted establishment of organisms on surfaces, impacts aquaculture facilities by decreasing the value of their products and causing expensive damages to their equipment. The biofouling tunicate *Didemnun vexillum* poses a notable threat to aquaculture given that it is an invasive species with strong competition abilities and is rapidly expanding its range. In this study, we seek to determine the impact of combining high-temperatures and freshwater treatments at different immersion times on *D. vexillum* as a method for controlling biofouling. We immersed *D. vexillum* in either freshwater or seawater at one of four different temperatures (12, 50, 70, and 90°C), for 60 or 120 seconds. We then analyzed the survival of the tunicate 3 weeks after treatment. We found that both 70°C and 90°C treatments successfully killed *D. vexillum* regardless of water type and immersion time. Therefore, to maximize the effectiveness of biofouling removal efforts while limiting the amount of time and energy used, we recommend aquaculture facilities should use 60 seconds 70°C seawater dips to control *D. vexillum* on their gear. Using this method to remove *D. vexillum* biofouling will help to decrease aquaculture gear damage, and reduce the spread of an invasive species.

## Libraries

```{r, message = FALSE}
library(tidyverse)
library(cowplot)
library(patchwork)
library(ggplot2)
library(here)
library(tidyr)
library(performance)
library(DHARMa)
library(fitdistrplus)
library(gamlss)
library(FSA)
library(goft)
library(MASS)
library(ordinal)
library(here)
```

## Analyses and Graphs
***

### Reading in Data

```{r}
here::here()
tunidata <- read_csv(here("Data", "tunicate_treatment_rawdata.csv"))
tunidata$temperature_c <- as.factor(tunidata$temperature_c)
tunidata$water_type <- as.factor(tunidata$water_type)
tunidata$exposure_time_s <- as.factor(tunidata$exposure_time_s)

```

### Change in mean RGB values

First Check for normality, p-value = 0.3433 normal distribution!

```{r}
shapiro.test(tunidata$X48hr_rgb)
```

Using a linear model for normal distribution - allows you to use random effects and nested effects

```{r, warning =FALSE}
tunirgb <- tunidata %>% 
  mutate(exposure_time_s = as.factor(exposure_time_s)) %>% 
  mutate(temperature_c = as.factor(temperature_c)) %>%
  mutate(change_rgb = X48hr_rgb-initial_rgb)

modrgb <- lm(X48hr_rgb ~ exposure_time_s + water_type + temperature_c +
               exposure_time_s*water_type*temperature_c + 
               (1|colony_id), family = gaussian, 
             data = tunirgb)
summary(modrgb)
```

Since model is not influenced by random effects, take this out of the model (stepAIC does not work with random effects) and then reduce model

```{r, warning =FALSE}
modrgb <- lm(X48hr_rgb ~ exposure_time_s + water_type + temperature_c + 
               exposure_time_s*water_type*temperature_c, family = gaussian, 
             data = tunirgb)
stepmodrgb <- stepAIC(modrgb, direction = "backward", trace = F)
formula(stepmodrgb)
```

stepAIC has outputted its final, reduced model. Final p values for model output

```{r}
newmodrgb <- lm(X48hr_rgb ~ water_type + temperature_c, data = tunirgb)
summary(newmodrgb)
```

Graph

```{r}

deltachangergb <- ggplot(tunirgb, aes(x = temperature_c, y = change_rgb, 
                                      fill = water_type))+
  labs(x = "Temperature (°C)", y = "Change in mean RGB", fill = "Water type")+
#labs(title = "Temperature (°C)")+
  scale_fill_manual(values = c("#0072B2","#E69F00"))+
  geom_jitter(position = position_dodge(width=0.75))+
  geom_boxplot(alpha = 0.8)+
  theme_bw()
deltachangergb

#ggsave("delta-change-rbg.jpg")
```

### Mold Cover

Displaying the data as a proportion of overall score with given mold cover scores

```{r}
tuni_stacked = tunidata %>%
  group_by(mold_cover, temperature_c, water_type, exposure_time_s) %>%
  summarise(frequency = n())%>%
  mutate(temperature_c=  as.factor(temperature_c))
```

Stacked bar graphs showing mold cover

```{r}
ggplot(tuni_stacked, aes(y = frequency, x = temperature_c, 
                         fill = as.factor(mold_cover))) +
  geom_bar(stat = "identity", position = "fill") +
  facet_grid(.~water_type) +
  labs(x ="Temperature (°C)", y = "Frequency")+
  scale_fill_brewer(palette = "Oranges", direction=1) +
  labs(fill = "Mold Cover") +
  theme_bw()

#ggsave("mold-cover.jpg")
```

Examining mold cover data to find the best distribution

```{r, warning =FALSE}
descdist(tunidata$mold_cover)
fit <- fitDist(mold_cover, data = tunidata, type = "realAll", try.gamlss = T)
fit$fits
```

Fitdist determined ST1 to be the best distribution, comparing it against a normal distribution to make sure AIC value is lower. AIC value of ST1 = -743.8 while AIC value of normal = 328

```{r, warning =FALSE}
mST1 <- histDist(tunidata$mold_cover, "ST1", density = T, 
                 main = "Skew t (Azzalini type 1)")
mNO <- histDist(tunidata$mold_cover, "NO", density = T, main = "Normal")
GAIC(mST1, mNO)
```

Now creating a full model, reduced model is the same as full model. No factors are significant

```{r, warning =FALSE}
tunirgb$colony_id <- as.factor(tunirgb$colony_id)
tunirgb$mold_cover <- as.factor(tunirgb$mold_cover)

ordinalmodfull <- clmm2(mold_cover ~ water_type + temperature_c + exposure_time_s +
                     exposure_time_s*water_type*temperature_c,
                     random = colony_id, data = tunirgb, Hess = TRUE)
summary(ordinalmodfull)

#temperature_c90 p=0.041566

```

### Change in weight after 48 hours

Data Visualization

```{r}
#Making new columns for changes in weight
tuni_data1 <- tunidata %>% 
  mutate(change_wet_weight_3wk = final_weight_g-post_acclimation_weight_g)
tuni_data1 <- tuni_data1 %>% 
  mutate(change_wet_weight_48 = `X48hr_weight`-post_acclimation_weight_g)

p2 <- ggplot(tuni_data1, aes(x = exposure_time_s, y = change_wet_weight_48, 
                             fill = water_type, ))+
  labs(x = "Immersion time (s)", y = "Change in wet weight (g)", 
       fill = "Water type")+
  geom_boxplot(alpha = 0.8)+
  scale_fill_manual(values = c("#0072B2","#E69F00"))+
  theme_bw()

p3 <- p2 + facet_grid(cols = vars(temperature_c))
p3
```

Testing for Normality - p-value = 0.01824 so not a normal distribution

```{r, warning =FALSE}
hist(tuni_data1$change_wet_weight_48)
shapiro.test(tuni_data1$change_wet_weight_48)
#p-value = 0.01824, not normal
descdist(tuni_data1$change_wet_weight_48)
#might be logistic
```

Distribution Fitting - used FitDist function - followed a logistic distribution

```{r, warning =FALSE}
fitDist(change_wet_weight_48, data=tuni_data1, type="realAll", try.gamlss = T)
#family=LO, logistic

mLOG_weight <- histDist(tuni_data1$change_wet_weight_48, "LO", density = T, 
                        main = "Logistic")

GAIC(mLOG_weight)
```

GAMLSS Model

```{r, warning =FALSE, results='hide'}
mod_weight <- gamlss(change_wet_weight_48 ~ water_type + temperature_c + 
                    exposure_time_s + water_type*temperature_c*exposure_time_s, 
                     family = LO, data = tuni_data1)
```

Model Selection - final formula for change

```{r, warning =FALSE}
stepmodweight48 <- stepGAIC(mod_weight, direction = "backward", trace = F)
summary(stepmodweight48)
#water_typeseawater:temperature_c50 p-value = 0.00922 significant

formula(stepmodweight48)
#change_wet_weight_48 ~ water_type + temperature_c + water_type:temperature_c
```

### Change in Weight over 3 Weeks (compare to post-acclimation)

Data Visualization

```{r}
p <- ggplot(tuni_data1, aes(x = exposure_time_s, y = change_wet_weight_3wk, 
                            fill = water_type))+
  labs(x = "Immersion time (s)", y = "Change in wet weight (g)", 
       fill = "Water type")+
#labs(title = "Temperature (°C)")+
  geom_boxplot(alpha = 0.8)+
  scale_fill_manual(values = c("#0072B2","#E69F00"))+
  theme_bw()

p4 <- p + facet_grid(cols = vars(temperature_c))

p3 + p4 + plot_layout(ncol=1)

#ggsave("change-in-wet-weight.jpg")
```

Testing for Normality

```{r, warning =FALSE}
hist(tuni_data1$change_wet_weight_3wk)
shapiro.test(tuni_data1$change_wet_weight_3wk)
#p-value = 0.7646 thus follows a normal distribution!
descdist(tuni_data1$change_wet_weight_3wk)
```

Distribution Fitting

```{r, warning =FALSE}
fitDist(change_wet_weight_3wk, data = tuni_data1, type = "realline", 
        try.gamlss = T)
#this also gives normal distribution
```

Linear Model

```{r, warning =FALSE}
mod_changeweight <- lm(change_wet_weight_3wk ~ 
                         exposure_time_s + water_type + 
                         temperature_c+ 
                         exposure_time_s*water_type*temperature_c, 
                       family = gaussian, data = tuni_data1)
```

Model Selection

```{r, warning =FALSE}
#Model Selection
step.mod_changeweight <- stepAIC(mod_changeweight, direction = "backward", 
                                 trace = F)
summary(step.mod_changeweight)
formula(step.mod_changeweight)

#change_wet_weight_3wk ~ water_type + temperature_c + water_type:temperature_c
```

Check Model

```{r, warning =FALSE}
check_model(step.mod_changeweight)
#check model when water_type:temperature_c was present and had major 
#collinearity issues, thus removed water_type:temperature and check model 
#again and collinearity issues were solved
```

New Model

```{r, warning =FALSE}
mod_changeweightv2 <- lm(change_wet_weight_3wk ~ water_type + temperature_c, 
                         family = gaussian, data = tuni_data1)
summary(mod_changeweightv2)
#significant p-value for temperature_c90 (p=value = 0.0315)
```

### Survival

Visualizing Data

```{r, message=FALSE}
#making a column for the propotion of samples that survived per treatment
tunisurv<- tunidata%>%
  group_by(temperature_c, water_type)%>%
  summarize(proportion_survival=sum(survival)/10, sd=sd(survival), total= n(), 
            SE = sd(survival)/sqrt(total))
# divided by 10 because each combination of 
#temperature and water type has 10 data points

```

```{r}
my_cols = c("freshwater" = "#0072B2", "seawater" = "#E69F00")

#plotting
pd <- position_dodge(width = 0.6)
ggplot(tunisurv, aes(x=temperature_c, y=proportion_survival, colour=water_type))+
  geom_point(aes(colour=water_type), position = pd, size=3)+
  xlab("Temperature (°C)")+
  ylab("Survival proportion")+
  geom_errorbar(aes(ymin=proportion_survival-SE, ymax=proportion_survival+SE, 
                    width=0.5), size=1.2, position = pd)+
  theme_bw()+labs(colour="Water type")+
    scale_color_manual(values = my_cols)+
  scale_fill_manual(values = my_cols)

pd
#ggsave("survivalproportion.jpg")
```

Testing for Normality

```{r}
shapiro.test(tunidata$survival)
#p-value = 2.047e-15
```

Checking Distribution

```{r, warning =FALSE}
fitDist(survival, data = tunidata, type = "binom", try.gamlss = T)
#family = BI (binomial)
```

GAMLSS model

```{r, warning =FALSE}
mod_survival <- gamlss(survival ~ water_type + temperature_c + 
                         exposure_time_s + 
                         water_type*temperature_c*exposure_time_s + 
                         random(as.factor(colony_id)), 
                       family = BI, data = tunidata)
summary(mod_survival)
```

Model Selection

```{r, warning =FALSE}
step.modsurvival <- stepGAIC(mod_survival, direction = "backward", trace = F)
summary(step.modsurvival)
formula(step.modsurvival)
#survival ~ temperature_c
```

Kruskal-Wallis Test

```{r}
kruskal.test(survival ~ temperature_c, data = tunidata)
#Kruskal-Wallis chi-squared = 26.171, df = 3, p-value = 8.781e-06
```

since temperature is the only explanatory variable - thus we can use Kruskal-Wallis Test to see p-values comparing temperatures to controls

```{r}

dunnTest(survival ~ temperature_c, data = tunidata)
#Comparison        Z      P.unadj        P.adj
#1    12 - 50 2.817892 4.834013e-03 1.933605e-02*
#2    12 - 70 4.428115 9.506008e-06 5.703605e-05*
#3    50 - 70 1.610224 1.073490e-01 3.220471e-01
#4    12 - 90 4.428115 9.506008e-06 4.753004e-05*
#5    50 - 90 1.610224 1.073490e-01 2.146980e-01
#6    70 - 90 0.000000 1.000000e+00 1.000000e+00

#p-value for 12-90 and 12-70 is significant!
```

### Attachment

Data Visualization

```{r, warning =FALSE, message = FALSE}
#making a column for proportion of samples attached to the container
tuniattach<- tunidata%>%
  group_by(temperature_c, water_type)%>%
  summarize(proportion_attached=sum(attachment)/10, sd=sd(attachment), 
            total= n(), SE = sd(attachment)/sqrt(total))
# divided by 10 because each temperature and water type combination has 10 data points
```

```{r}

#plotting
ggplot(tuniattach, aes(x=temperature_c, y=proportion_attached, colour=water_type))+
  geom_point(aes(colour=water_type), position=pd, size=3)+
  xlab("Temperature (°C)")+
  ylab("Attachment proportion")+
  geom_errorbar(aes(ymin=proportion_attached-SE, ymax=proportion_attached+SE,
                    width=.5), size=1.2, position=pd)+
  theme_bw()+labs(colour="Water type")+
  scale_color_manual(values = my_cols)+
  scale_fill_manual(values = my_cols)




#ggsave("attachmentproportion.jpg")
```

Distribution Fitting

```{r, warning =FALSE}
fitDist(attachment, data = tuni_data1, type = "binom", try.gamlss = T)
#family = BI (binomial)
```

GAMLSS Model

```{r, warning =FALSE}
mod <- gamlss(attachment ~ water_type + temperature_c + 
                exposure_time_s + 
                water_type*temperature_c*exposure_time_s, 
              family = BI, data = tuni_data1)
summary(mod)
```

Model Selection

```{r, warning =FALSE}
step.mod <- stepAIC(mod, direction = "backward", trace = F)
summary(step.mod)
formula(step.mod)
#attachment ~ temperature_c
```

Kruskie

```{r}
kruskal.test(attachment ~ temperature_c, data = tunidata)
#Kruskal-Wallis chi-squared = 37.783, df = 3, p-value = 3.142e-08
```

since temperature is the only explanatory variable - thus we can use Kruskal-Wallis Test to see p-values comparing temperatures to controls

```{r, warning =FALSE}
dunnTest(attachment ~ temperature_c, data = tuni_data1)
#Comparison        Z      P.unadj        P.adj
#1    12 - 50 5.018805 5.199384e-07 3.119630e-06*
#2    12 - 70 5.018805 5.199384e-07 2.599692e-06*
#3    50 - 70 0.000000 1.000000e+00 1.000000e+00
#4    12 - 90 5.018805 5.199384e-07 2.079753e-06*
#5    50 - 90 0.000000 1.000000e+00 1.000000e+00
#6    70 - 90 0.000000 1.000000e+00 1.000000e+00

#p-value for 12-90, 12-70 and 12-50 are significant
```
